---
title: "Sensitivity analysis model parameters noLight"
author: "Daniel Gschwentner"
date: "2024-07-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(include = F)
```

## Note for Carly

I changed the code to show differences as percent instead of relative ratios. I've subtracted 100% from each simulation to normalize/show the percent different (i.e. a value below 0% means decrease relative to baseline, a value \>0% means an increase). You'll notice that the % change is quite large for a few parameters (e.g. \>1000% increase relative to baseline). I think this messes with the visualization as the color scale becomes misleading: basically, it is impossible to discern small changes because the larger values stretch the color palette. We can probably fix this with some clever coding of the color scheme (e.g. use a categorical color scheme) but I thought the relative changes helped circumvent this issue.

### Notes

Somehow everything revolves around N:P of 16 even when using average phytoplankton stoichiometry... there seem to be shifts in sensitivity depending on whether N or P limitation prevails.

Much greater changes to baseline in Michaelis-Menten model compared to Droop model –\> luxury uptake allows for decoupling of growth and resource supply –\> makes model less sensitive? provides buffering capacity?!

## Overview

Code to perform sensitivity analysis of algal physiology parameters for Michaelis-Menten and Droop models.

Models are run across TN and TP inflow gradients of 100 - 1000 ug L^-1^ and 10 - 500 ug L^-1^ respectively to establish baselines (500 timesteps). Light is held constant at I~0~ = 400 (umoles m^-2^ s^-1^ ). Algal physiology parameters are then varied by +/-10% to evaluate how sensitive C-acquisition is to changes in physiology. **Note, that I haven't performed a sensitivity analysis for umax or K~Light~ as these variables use fundamentally different units than the nutrient/stoichiometry parameters. When included umax appears to be the most influential variable and obscures variation in the other variables. Can easily include umax in the future if folks would like**. In addition, sensitivity of cellular N/P to changes in algal physiology is determined for the Droop model.

## Setup

```{r packages, echo = F, message=F, warning=F, results = "hide"}

pck <- c("deSolve", "tidyverse", "cowplot", "ggsci")
lapply(pck, require, character.only = T)
theme_set(theme_cowplot() + theme(legend.justification = "center"))
# load ODE models 
# saved in external files for convenience
# Michalis-Menten model
source("models/mich_single_noLight.R") # michaelis menten kinetics, one algae species; model 1 in Carly's framework
# Droop model
source("models/droop_single_noLight.R") # droop model, one algae species; model 3 in Carly's framework
```

For the models I am using the average parameter values from the table below. These parameters are from [Edwards et al., 2016](https://figshare.com/articles/dataset/Data_Paper_Data_Paper/3562857?file=5635515).

```{r load traits, echo = F, include=T}
# trait data
# median values
traits <- read_csv("data4input/phyto_traits4models_21June2024.csv")[,-1]
knitr::kable(traits, caption = "Trait parameters")

```

### Model setup

```{r create grid of nutrient inflow scenarios}

# timesteps
times <- 1:500
# loads
loads <- expand.grid(Pin = seq(10, 500,50 ),
                                Nin = seq(10,1500 , 100))

```

```{r michaelis-menten model setup}

params.mich <- c(
  # lake parameters
  SA= 1,		# lake surface area in km2
  zmix = 2, # lake mixing depth in m
  Pin = NA, # P inflow concentration in mg P m^-3
  Nin = NA, # N inflow concentration in mg N m^-3
  
  # light parameters
  I0 = 400, # incident light
  kBg= 0.1,		# background light attenuation (m-1)/(g C m-3); 0.1-5.6 from Jager & Diehl
  kA=0.0003, # algal light attenuation (m-1)/(mg C m-3); 0.0003 from Jager & Diehl

  # algae physiology parameters
  umax1 = traits[12, "average"],
  lA=0.1,			# mortality rate day-1
  v=0.1,			# m d-1; sinking loss of algae
  KLight = traits[11, "average"], # light half sat constant 
  KP1 = traits[2, "average"] * 1000, # phosphorus half sat constant in mg P m^-3 f
  QP1 = traits[6, "average"], # algae cell P quota in mg P mg^-1 C^-1
  KN1 = traits[1, "average"] * 1000, # nitrogen half sat constant in mg N m^-3 
  QN1 = traits[5, "average"] # algae cell N quota in mg N mg^-1 C^-1 
)
names(params.mich) <- c("SA", "zmix", "Pin", "Nin", "I0", 
                        "kBg", "kA", "umax1", "lA", "v", 
                        "KLight", "KP1",  "QP1", "KN1",  "QN1")
names(params.mich)
params.mich <- unlist(params.mich)

```

```{r droop model parameter setup}
params.droop <- c(
  # lake parameters
  SA= 1,		# lake surface area in km2
  zmix = 2, # lake mixing depth in m
  Pin = NA, # P inflow concentration in mg P m^-3
  Nin = NA, # N inflow concentration in mg N m^-3

  # light parameters
  I0 = 400, # surface light
  kBg= 0.1,		# background light attenuation (m-1)/(g C m-3); 0.1-5.6 from Jager & Diehl
  kA=0.0003, # algal light attenuation (m-1)/(mg C m-3); 0.0003 from Jager & Diehl
 
  # algae physiology parameters
  umax1 = traits[12, "average"],
  lA=0.1,			# mortality rate day-1
  v= 0.1,			# m d-1; sinking loss of algae
  KLight = traits[11, "average"], # light half sat constant 
  KP1 = traits[2, "average"] * 1000, # phosphorus half sat constant in mg P m^-3 f
  minQP1 = traits[6, "average"], # algae cell P quota in mg P mg^-1 C^-1 f
  upP1 = traits[4, "average"], # max uptake rate P per day in mg P mg C^-1 day^-1 
  KN1 = traits[1, "average"] * 1000, # nitrogen half sat constant in mg N m^-3 
  minQN1 = traits[5, "average"], # algae cell N quota in mg N mg^-1 C^-1 f
  upN1 = traits[3, "average"] # max uptake rate N per day in mg N mg C^-1 day^-1 
)



names(params.droop) <- c("SA", "zmix", "Pin", "Nin", "I0", "kBg", "kA",
                             "umax1", "lA", "v", "KLight", 
                             "KP1", "minQP1", "upP1", "KN1",
                             "minQN1", "upN1")
names(params.droop)
params.droop <- unlist(params.droop)
```

## Run baseline models

### Michaelis Menten model

```{r michaelis menten baselines}

baselines.mich <- lapply(1:nrow(loads), function(i) {
  params.mich["Pin"] = loads[i, "Pin"]
  params.mich["Nin"] = loads[i, "Nin"]
  y <- c("A1" = 100,
         "P" = loads[i, "Pin"],
         "N" = loads[i, "Nin"])
  
  run <- ode(y, times, parms = params.mich, func = mich.single)
  return(run[max(times),])
  
})

baselines.mich <- do.call(rbind, baselines.mich)
baselines.mich <- as_data_frame(baselines.mich)
baselines.mich$Pin <- loads$Pin
baselines.mich$Nin <- loads$Nin

```

### Droop models

```{r}
baselines.droop <- lapply(1:nrow(loads), function(i) {
  params.droop["Pin"] = loads[i, "Pin"]
  params.droop["Nin"] = loads[i, "Nin"]
  y <- c("A1" = 100,
         "P" = loads[i, "Pin"],
         "N" = loads[i, "Nin"],
         "QP1" = 0.015,
         "QN1" = 0.1)
  
  run <- ode(y, times, parms = params.droop, func = droop.single)
  return(run[max(times),])
  
})

baselines.droop <- do.call(rbind, baselines.droop)
baselines.droop <- as_data_frame(baselines.droop)
baselines.droop$Pin <- loads$Pin
baselines.droop$Nin <- loads$Nin


```

## Sensitivity analysis

### Michaelis-Menten models

Increase and decrease algal physiology parameters by +/-10% and store the output. Calculate the change relative to baseline conditions across TN and TP loads.

Code chunk below varies parameters by +10%

```{r michales menten pt1}
# iterate over parameters
sim.mich.1 <- lapply(list("KP1","QP1", "KN1", "QN1"), function(f) {
  params.mich[f] <- params.mich[f] * 1.1
  # iterate over loads
  sim.1 <- lapply(1:nrow(loads), function(i) {
    params.mich["Pin"] = loads[i, "Pin"]
    params.mich["Nin"] = loads[i, "Nin"]
    y <- c("A1" = 100, "P" = loads[i, "Pin"], "N" = loads[i, "Nin"])
    run <- ode(y, times, parms = params.mich, func = mich.single)
    return(run[max(times),])
  })
  # add meta-data
  sim.1 <- do.call(rbind, sim.1)
  sim.1 <- as_data_frame(sim.1)
  sim.1$Pin <- loads$Pin
  sim.1$Nin <- loads$Nin
  # add info on the parameter that was varied
  sim.1$param <- f
  return(sim.1)
})
sim.mich.1 <- do.call(rbind, sim.mich.1)
sim.mich.1$scenario <- "+10%"


```

Code chunk below varies parameters by -10%\
(there is probably a more elegant way to do this but this simple approach works for now.

```{r michales menten pt2}
# iterate over parameters
sim.mich.2 <- lapply(list("KP1", "QP1", "KN1", "QN1"), function(f) {
  params.mich[f] <- params.mich[f] * 0.9
  # iterate over loads
  sim.1 <- lapply(1:nrow(loads), function(i) {
    params.mich["Pin"] = loads[i, "Pin"]
    params.mich["Nin"] = loads[i, "Nin"]
    y <- c("A1" = 100, "P" = loads[i, "Pin"], "N" = loads[i, "Nin"])
    run <- ode(y, times, parms = params.mich, func = mich.single)
    return(run[max(times),])
  })
  # add meta-data
  sim.1 <- do.call(rbind, sim.1)
  sim.1 <- as_data_frame(sim.1)
  sim.1$Pin <- loads$Pin
  sim.1$Nin <- loads$Nin
  sim.1$param <- f
  return(sim.1)
})
sim.mich.2 <- do.call(rbind, sim.mich.2)
sim.mich.2$scenario <- "-10%"


```

The code chunk below re-formats the data and merges the simulations with the baseline runs. I've calculated the relative change from the simulation to the baseline for a given TN and TP load as

$$
\Delta_{biomass\%} = \frac{sensitivity_i * 100\%}{baseline}
$$

where $\Delta_{biomass}$ is the relative change in biomass, baseline is the baseline biomass, and sensitivity~i~ is the biomass for a given parameter in the sensitivity analysis.

In addition, I calculated which parameter results in the maximum change in biomass relative to the baseline. It is important to note that this doesn't tell us the direction of change just the magnitude of change.

```{r merge data sets and fix names}

# merge scenarios
sim.mich <- rbind(sim.mich.1, sim.mich.2)
colnames(sim.mich) <- c("time", "A1.sim", "P.sim", "N.sim",
                        "Pin", "Nin", "param", "scenario")
sim.mich <- merge(sim.mich, baselines.mich, 
                  by = c("Pin", "Nin"))
# calculate percent change relative to baseline
sim.mich$change2baseline <- sim.mich$A1.sim * 100/sim.mich$A1 - 100

# determine min/max change
# note that this code doesn't tell us about direction of change, only the magnitude
sim.mich.metrics <- sim.mich %>%
  mutate(change2baseline2 =  abs(change2baseline)) %>% 
  pivot_wider(id_cols = c(Pin, Nin, scenario),
              names_from = param, 
              values_from = change2baseline2) 
colnames(sim.mich.metrics)
sim.mich.metrics[,-c(1, 2, 3)] <- apply(sim.mich.metrics[,-c(1, 2, 3)], 2, abs)
sim.mich.metrics$max.change <- colnames(sim.mich.metrics[-c(1,2,3)])[apply(sim.mich.metrics[,-c(1,2,3)], 1, which.max)]


```

### Droop models

Increase and decrease algal physiology parameters by +/-10% and store the output. Calculate the change relative to baseline conditions across TN and TP loads.

```{r droop pt 1}
# iterate over parameters
sim.droop.1 <- lapply(list("KP1","KN1", "upP1", "upN1", 
                           "minQP1", "minQN1"), function(f) {
  params.droop[f] <- params.droop[f] * 1.1
  # iterate over loads
  sim.1 <- lapply(1:nrow(loads), function(i) {
    params.droop["Pin"] = loads[i, "Pin"]
    params.droop["Nin"] = loads[i, "Nin"]
    y <- c("A1" = 100, "P" = loads[i, "Pin"], "N" = loads[i, "Nin"],
           "QP1" = 0.015,"QN1" = 0.1)
    run <- ode(y, times, parms = params.droop, func = droop.single)
    return(run[max(times),])
  })
  # add meta-data
  sim.1 <- do.call(rbind, sim.1)
  sim.1 <- as_data_frame(sim.1)
  sim.1$Pin <- loads$Pin
  sim.1$Nin <- loads$Nin
  # add info on the parameter that was varied
  sim.1$param <- f
  return(sim.1)
})
sim.droop.1 <- do.call(rbind, sim.droop.1)
sim.droop.1$scenario <- "+10%"

```

```{r droop pt2}

# iterate over parameters
sim.droop.2 <- lapply(list("KP1","KN1", "upP1", "upN1", 
                           "minQP1", "minQN1"), function(f) {
  params.droop[f] <- params.droop[f] * 0.9
  # iterate over loads
  sim.1 <- lapply(1:nrow(loads), function(i) {
    params.droop["Pin"] = loads[i, "Pin"]
    params.droop["Nin"] = loads[i, "Nin"]
    y <- c("A1" = 100, "P" = loads[i, "Pin"], "N" = loads[i, "Nin"],
           "QP1" = 0.015,"QN1" = 0.1)
    run <- ode(y, times, parms = params.droop, func = droop.single)
    return(run[max(times),])
  })
  # add meta-data
  sim.1 <- do.call(rbind, sim.1)
  sim.1 <- as_data_frame(sim.1)
  sim.1$Pin <- loads$Pin
  sim.1$Nin <- loads$Nin
  # add info on the parameter that was varied
  sim.1$param <- f
  return(sim.1)
})
sim.droop.2 <- do.call(rbind, sim.droop.2)
sim.droop.2$scenario <- "-10%"

```

The code chunk below re-formats the data and merges the simulations with the baseline runs. I've calculated the relative change from the simulation to the baseline for a given TN and TP load as

$$ \Delta_{biomass\%} = \frac{sensitivity_i * 100%}{baseline} $$

where $\Delta_{biomass}$ is the relative change in biomass, baseline is the baseline biomass, and sensitivity~i~ is the biomass for a given parameter in the sensitivity analysis.

In addition, I calculated which parameter results in the maximum change in biomass relative to the baseline. It is important to note that this doesn't tell us the direction of change just the magnitude of change.

```{r merge data sets and fix names droop model}
# merge scenarios
sim.droop <- rbind(sim.droop.1, sim.droop.2)
colnames(sim.droop) <- c("time", "A1.sim", "P.sim", "N.sim",
                        "QP.sim", "QN.sim", "Pin", "Nin", 
                        "param", "scenario")
sim.droop <- merge(sim.droop, baselines.droop, 
                  by = c("Pin", "Nin"))
# calculate percent change relative to baseline
# for algal biomass
sim.droop$change2baseline.A <- sim.droop$A1.sim * 100/sim.droop$A1 - 100
# for cell P content
sim.droop$change2baseline.QP <- sim.droop$QP.sim * 100/sim.droop$QP1 - 100
# for cell N content
sim.droop$change2baseline.QN <- sim.droop$QN.sim * 100/sim.droop$QN1 - 100

# determine min/max change
# note that this code doesn't tell us about direction of change, only the magnitude
sim.droop.metrics.A <- sim.droop %>% 
  mutate(change2baseline2 =  abs(change2baseline.A)) %>% 
  pivot_wider(id_cols = c(Pin, Nin, scenario),
              names_from = param, 
              values_from = change2baseline2)
colnames(sim.droop.metrics.A)
sim.droop.metrics.A[,-c(1, 2, 3)] <- apply(sim.droop.metrics.A[,-c(1, 2, 3)], 2, abs)
sim.droop.metrics.A$max.change <- colnames(sim.droop.metrics.A[-c(1,2,3)])[apply(sim.droop.metrics.A[,-c(1,2,3)], 1, which.max)]

# for QP
sim.droop.metrics.QP <- sim.droop %>% 
  mutate(change2baseline2 =  abs(change2baseline.QP)) %>% 
  pivot_wider(id_cols = c(Pin, Nin, scenario),
              names_from = param, 
              values_from = change2baseline2)
colnames(sim.droop.metrics.QP)
sim.droop.metrics.QP[,-c(1, 2, 3)] <- apply(sim.droop.metrics.QP[,-c(1, 2, 3)], 2, abs)
sim.droop.metrics.QP$max.change <- colnames(sim.droop.metrics.QP[-c(1,2,3)])[apply(sim.droop.metrics.QP[,-c(1,2,3)], 1, which.max)]

# for QN
sim.droop.metrics.QN <- sim.droop %>%
  mutate(change2baseline2 =  abs(change2baseline.QN)) %>%
  pivot_wider(id_cols = c(Pin, Nin, scenario),
              names_from = param, 
              values_from = change2baseline2)
colnames(sim.droop.metrics.QN)
sim.droop.metrics.QN[,-c(1, 2, 3)] <- apply(sim.droop.metrics.QN[,-c(1, 2, 3)], 2, abs)
sim.droop.metrics.QN$max.change <- colnames(sim.droop.metrics.QN[-c(1,2,3)])[apply(sim.droop.metrics.QN[,-c(1,2,3)], 1, which.max)]

```

## Summary

### Michaelis-Menten

```{r}

mich.menten.1 <- sim.mich %>%
  ggplot(aes(Pin, Nin, fill = change2baseline)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_grid(scenario ~ param) + 
  #scale_fill_viridis_c(option = "H") + 
  scale_fill_viridis_b(breaks = c(
    -100, -50, 0, 10, 50, 100, 500, 1000
  ), option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "% change")

mich.menten.2 <- sim.mich.metrics %>%
  ggplot(aes(Pin, Nin, fill = max.change)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_wrap(scenario~.,ncol = 2) + 
  scale_fill_viridis_d(option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "max. change")

mich.menten.3 <- sim.mich %>%
    ggplot(aes(param, change2baseline)) + 
    geom_violin() + 
    facet_wrap(scenario~., scales = "free") + 
    labs(x = NULL, y = "relative change")


```


```{r, dpi=500, fig.width=7, fig.height = 3, include=T}
mich.menten.1 

```


```{r, dpi=500, fig.width=5, fig.height=3, fig.width = 8, include=T}
mich.menten.2

```

```{r, dpi = 500, fig.width=7, fig.height=3}
#mich.menten.3

```

#### Conclusion for Michaelis-Menten models

### Droop models

```{r }
droop.A <- sim.droop %>%
  ggplot(aes(Pin, Nin, fill = change2baseline.A)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_grid(scenario ~ param) + 
  #scale_fill_viridis_c(option = "H") + 
    scale_fill_viridis_b(breaks = c(
    -100, -50, 0, 10, 50, 100, 500, 1000
  ), option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "% change")

droop.QP <- sim.droop %>%
  ggplot(aes(Pin, Nin, fill = change2baseline.QP)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_grid(scenario ~ param) + 
  #scale_fill_viridis_c(option = "H") + 
    scale_fill_viridis_b(breaks = c(
    -100, -50, 0, 10, 50, 100, 500, 1000
  ), option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "% change")

droop.QN <- sim.droop %>%
  ggplot(aes(Pin, Nin, fill = change2baseline.QN)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_grid(scenario ~ param) + 
  #scale_fill_viridis_c(option = "H") + 
    scale_fill_viridis_b(breaks = c(
    -100, -50, 0, 10, 50, 100, 500, 1000
  ), option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "% change")

droop.max.A <- sim.droop.metrics.A %>%
  ggplot(aes(Pin, Nin, fill = max.change)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_wrap(scenario~., ncol = 2) + 
  scale_fill_viridis_d(option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "max. change")

droop.max.QP <- sim.droop.metrics.QP %>%
  ggplot(aes(Pin, Nin, fill = max.change)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_grid(scenario~.) + 
  facet_wrap(scenario~., ncol = 2) + 
  scale_fill_viridis_d(option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "max. change")

droop.max.QN <- sim.droop.metrics.QN %>%
  ggplot(aes(Pin, Nin, fill = max.change)) + 
  geom_raster() + 
  geom_abline(intercept = 0, slope = (16 * 14.007)/30.974,
              lwd = 1, lty = "dashed", col = "white") + 
  facet_wrap(scenario~., ncol = 2) + 
  scale_fill_viridis_d(option = "H") + 
  scale_x_continuous(breaks = c(10, 250, 500)) + 
  labs(x = "P in ppm", y = "N in ppm", fill = "max. change")

```

#### Algal biomass



```{r dpi=500, fig.width=8, fig.height = 3, include=T}

droop.A

```


```{r dpi = 500, fig.width=7, fig.height=3, include=T}

droop.max.A

```

#### Cellular P content


```{r dpi=500, fig.width=8, fig.height = 3, include=T}

droop.QP
```


```{r dpi = 500, fig.width=7, fig.height=3, include=T}

droop.max.QP

```

#### Cellular N

```{r dpi=500, fig.width=8, fig.height = 3, include=T}

droop.QN

```


```{r dpi = 500, fig.width=7, fig.height=3, include=T}

droop.max.QN
```
